How Significant was WW2 in the Development of the Computer?

### 2017-01-29 Draft 2/Section 1 - Lambda Calculus.tex
Added Turing Machines section
(empty); Fixed numbering
Corrections on Sec1
Quick corrections to Factorial function

\begin{document}
\pagenumbering{roman}

@@

Multiplication can also be calculated by defining the function \textbf{M}:
\begin{equation*}
\bm{\mathrm{M}} = \lambda wyx.w(yx)
\end{equation*}
\begin{gather*}
\begin{aligned}
\bm{\mathrm{M22}} &= (\lambda wyx.w(yx))(\lambda ab.a(a(b)))(\lambda sz.s(s(z)))\\
&= \lambda x.(\lambda ab.a(a(b)))[(\lambda sz.s(s(z)))x]\\
&= \lambda x.(\lambda ab.a(a(b)))[\lambda z.x(x(z))]\\
&= \lambda x.(\lambda b.[\lambda z.x(x(z))]([\lambda z.x(x(z))]b))\\
&= \lambda x.(\lambda b.[\lambda z.x(x(z))][x(x(b))])\\
&= \lambda x.(\lambda b.x(x(x(x(b)))))\\
&= \lambda xb.x(x(x(x(b))))\\
&= \bm{\mathrm{4}}
\end{aligned}
\end{gather*}
We should rearrange our earlier addition method into an 'addition' function of the same format which takes two arguments:
\begin{equation*}
\bm{\mathrm{A}} = \lambda ab.aSb
\end{equation*}

@@

These concepts are the base of conditional logic in computation, and in $\lambda$-calculus operate like \textit{if} statements found in most programming languages. For such operations, the concept of a 'pair' exists:
\begin{equation*}
\lambda x.xab
\end{equation*}
where $a$ and $b$ are functions. For a choice of an element from the pair, the argument of either \textbf{T} or \textbf{F} is provided:
\begin{gather*}
\begin{aligned}
(\lambda x.xab)T &= a\\
(\lambda x.xab)F &= b
\end{aligned}
\end{gather*}

@@

Firstly, the function transforms the pair (n, n-1) (the argument $p$) into the pair (n+1, n) (the new pair's elements are an increment and a copy of the first value respectively):
\begin{equation*}
\bm{\psi} = \lambda pc.c(S(pT))(pT)
\end{equation*}
To find the predecessor to $n$ therefore, we apply this function $n$ times to the pair (0, 0) and the pair (n, n-1) is formed, after which the second element can easily be selected.
\begin{equation*}
\bm{\mathrm{P}} = \lambda n.n \bm{\psi} (\lambda x.00) F 
\end{equation*}
\begin{equation*}
\begin{aligned}
\bm{\mathrm{P3}} &= (\lambda n.n [\lambda pc.c(\bm{\mathrm{S}}(p\bm{\mathrm{T}}))(p\bm{\mathrm{T}})] [\lambda x.\bm{\mathrm{00}}] \bm{\mathrm{F}}) [\lambda sz.s(s(s(z)))]\\
&= ([\lambda sz.s(s(s(z)))] [\lambda pc.c(\bm{\mathrm{S}}(p\bm{\mathrm{T}}))(p\bm{\mathrm{T}})] [\lambda x.\bm{\mathrm{00}}]) \bm{\mathrm{F}}\\
&= ([\lambda sz.s(s(s(z)))] [\lambda pc.c(\bm{\mathrm{S}}(p\bm{\mathrm{T}}))(p\bm{\mathrm{T}})] [\lambda x.\bm{\mathrm{00}}]) \bm{\mathrm{F}}\\
&= [\lambda pc.c(\bm{\mathrm{S}}(p\bm{\mathrm{T}}))(p\bm{\mathrm{T}})]([\lambda pc.c(\bm{\mathrm{S}}(p\bm{\mathrm{T}}))(p\bm{\mathrm{T}})]([\lambda pc.c(\bm{\mathrm{S}}(p\bm{\mathrm{T}}))(p\bm{\mathrm{T}})][\lambda x.\bm{\mathrm{00}}]))\bm{\mathrm{F}}\\
&= [\lambda pc.c(\bm{\mathrm{S}}(p\bm{\mathrm{T}}))(p\bm{\mathrm{T}})]([\lambda pc.c(\bm{\mathrm{S}}(p\bm{\mathrm{T}}))(p\bm{\mathrm{T}})](\lambda c.c(\bm{\mathrm{S}}[\lambda x.\bm{\mathrm{00}}]\bm{\mathrm{T}})([\lambda x.\bm{\mathrm{00}}]\bm{\mathrm{T}})))\bm{\mathrm{F}}\\
&= [\lambda pc.c(\bm{\mathrm{S}}(p\bm{\mathrm{T}}))(p\bm{\mathrm{T}})]([\lambda pc.c(\bm{\mathrm{S}}(p\bm{\mathrm{T}}))(p\bm{\mathrm{T}})](\lambda c.c\bm{\mathrm{10}}))\bm{\mathrm{F}}\\
&= [\lambda pc.c(\bm{\mathrm{S}}(p\bm{\mathrm{T}}))(p\bm{\mathrm{T}})][\lambda c.c\bm{\mathrm{21}}]\bm{\mathrm{F}}\\
&= [\lambda c.c\bm{\mathrm{32}} ])\bm{\mathrm{F}}\\
&= \bm{\mathrm{2}}
\end{aligned}
\end{equation*}

Of course now the subtraction function can be found (using \textbf{D} for difference):
\begin{equation*}
\bm{\mathrm{D}} = \lambda ab.bPa
\end{equation*}
Bearing in mind the predecessor of \textbf{0} is \textbf{0}, this will produce \textbf{0} if $b>a$.
\subsubsection{Further Conditional Logic}

@@

\begin{equation*}
\bm{\mathrm{R}} = \lambda y.(\lambda x.y(xx))(\lambda x.y(xx)
\end{equation*}
\begin{equation*}
\begin{aligned}
\bm{\mathrm{R}}a &= (\lambda x.a(xx))[\lambda x.a(xx)]\\
&= a([\lambda x.a(xx)][\lambda x.a(xx)]\\
&= a(\bm{\mathrm{R}}a)
\end{aligned}
\end{equation*}
Therefore, the \textbf{R} function applied to any function $a$ applies the function $a$ to the function \textbf{R} applied to $a$ - if substitution continues, an infinite series of nested $a$ functions is created; an infinite result is produced from a finite configuration. Of course we can create finites as well by halting the function - an example of this would be the factorial function, where \textbf{!}a = \textbf{M}a(\textbf{!}[\textbf{P}a]) and \textbf{!0} = \textbf{1}. Firstly, a recursive definition:
\begin{equation*}
\begin{aligned}
\bm{\mathrm{\xi}} &= (\lambda fa.\bm{\mathrm{Z}}a\bm{\mathrm{1}}[\bm{\mathrm{M}}a(f[\bm{\mathrm{P}}a])])\\
\bm{!} &= \bm{\xi}!
\end{aligned}
\end{equation*}
i.e. the factorial function is the $\bm{\xi}$ function with itself as the argument $f$. Unfortunately such a definition is illegal under $\lambda$-Calculus as strictly no functions have names - this is where the function \textbf{R} comes into play. To find the factorial of \textbf{3}:
\begin{equation*}
\begin{aligned}
\bm{\mathrm{R\xi 3}} &= \bm{\mathrm{\xi}}(\bm{\mathrm{R\xi}})\bm{\mathrm{3}}\\
&= \bm{\mathrm{Z30}}[\bm{\mathrm{M3}}([\bm{\mathrm{R\xi}}][\bm{\mathrm{P3}}])\\
&= \bm{\mathrm{M3}}(\bm{\mathrm{R\xi}}\bm{2})]\\
&= \bm{\mathrm{M3}}(\bm{\mathrm{M2}}(\bm{\mathrm{M11}}))\\
&= \bm{\mathrm{6}}
\end{aligned}
\end{equation*}
The recursive identity \textbf{R} allows for \textbf{R}$\bm{\xi}$ to be applied as the first argument to the $\bm{\xi}$ function - as a result, the procedure is repeated for \textbf{2} and \textbf{1}, before stopping at \textbf{0}. To simplify, the factorial function:
\begin{equation*}
\bm{!} = \lambda x.\bm{\mathrm{R\xi}}x
\end{equation*}

>>>>

\begin{document}

@@

Multiplication can also be calculated by defining the function \textbf{M}:
\begin{equation*}
\bm{\mathrm{M}} = \lambda wyx.w(yx)
\end{equation*}
\begin{gather*}
\begin{aligned}
\bm{\mathrm{M22}} &= (\lambda wyx.w(yx))(\lambda ab.a(a(b)))(\lambda sz.s(s(z)))\\
&= \lambda x.(\lambda ab.a(a(b)))[(\lambda sz.s(s(z)))x]\\
&= \lambda x.(\lambda ab.a(a(b)))[\lambda z.x(x(z))]\\
&= \lambda x.(\lambda b.[\lambda z.x(x(z))]([\lambda z.x(x(z))]b))\\
&= \lambda x.(\lambda b.[\lambda z.x(x(z))][x(x(b))])\\
&= \lambda x.(\lambda b.x(x(x(x(b)))))\\
&= \lambda xb.x(x(x(x(b))))\\
&= \bm{\mathrm{4}}
\end{aligned}
\end{gather*}
We should rearrange our earlier addition method into an 'addition' function of the same format which takes two arguments:
\begin{equation*}
\bm{\mathrm{A}} = \lambda ab.a\bm{\mathrm{S}}b
\end{equation*}

@@

These concepts are the base of conditional logic in computation, and in $\lambda$-calculus operate like \textit{if} statements found in most programming languages. For such operations, the concept of a 'pair' exists:
\begin{equation*}
\lambda x.xab
\end{equation*}
where $a$ and $b$ are functions. For a choice of an element from the pair, the argument of either \textbf{T} or \textbf{F} is provided:
\begin{gather*}
\begin{aligned}
(\lambda x.xab)\bm{\mathrm{T}} &= a\\
(\lambda x.xab)\bm{\mathrm{F}} &= b
\end{aligned}
\end{gather*}

@@

Firstly, the function transforms the pair (n, n-1) (the argument $p$) into the pair (n+1, n) (the new pair's elements are an increment and a copy of the first value respectively):
\begin{equation*}
\bm{\psi} = \lambda pc.c(\bm{\mathrm{S}}(p\bm{\mathrm{T}}))(p\bm{\mathrm{T}})
\end{equation*}
To find the predecessor to $n$ therefore, we apply this function $n$ times to the pair (0, 0) and the pair (n, n-1) is formed, after which the second element can easily be selected.
\begin{equation*}
\bm{\mathrm{P}} = \lambda n.n \bm{\psi} (\lambda x.\bm{\mathrm{00}}) \bm{\mathrm{F}} 
\end{equation*}
\begin{equation*}
\begin{aligned}
\bm{\mathrm{P3}} &= (\lambda n.n [\lambda pc.c(\bm{\mathrm{S}}(p\bm{\mathrm{T}}))(p\bm{\mathrm{T}})] [\lambda x.\bm{\mathrm{00}}] \bm{\mathrm{F}}) [\lambda sz.s(s(s(z)))]\\
&= ([\lambda sz.s(s(s(z)))] [\lambda pc.c(\bm{\mathrm{S}}(p\bm{\mathrm{T}}))(p\bm{\mathrm{T}})] [\lambda x.\bm{\mathrm{00}}]) \bm{\mathrm{F}}\\
&= ([\lambda sz.s(s(s(z)))] [\lambda pc.c(\bm{\mathrm{S}}(p\bm{\mathrm{T}}))(p\bm{\mathrm{T}})] [\lambda x.\bm{\mathrm{00}}]) \bm{\mathrm{F}}\\
&= [\lambda pc.c(\bm{\mathrm{S}}(p\bm{\mathrm{T}}))(p\bm{\mathrm{T}})]([\lambda pc.c(\bm{\mathrm{S}}(p\bm{\mathrm{T}}))(p\bm{\mathrm{T}})]([\lambda pc.c(\bm{\mathrm{S}}(p\bm{\mathrm{T}}))(p\bm{\mathrm{T}})][\lambda x.\bm{\mathrm{00}}]))\bm{\mathrm{F}}\\
&= [\lambda pc.c(\bm{\mathrm{S}}(p\bm{\mathrm{T}}))(p\bm{\mathrm{T}})]([\lambda pc.c(\bm{\mathrm{S}}(p\bm{\mathrm{T}}))(p\bm{\mathrm{T}})](\lambda c.c(\bm{\mathrm{S}}[\lambda x.\bm{\mathrm{00}}]\bm{\mathrm{T}})([\lambda x.\bm{\mathrm{00}}]\bm{\mathrm{T}})))\bm{\mathrm{F}}\\
&= [\lambda pc.c(\bm{\mathrm{S}}(p\bm{\mathrm{T}}))(p\bm{\mathrm{T}})]([\lambda pc.c(\bm{\mathrm{S}}(p\bm{\mathrm{T}}))(p\bm{\mathrm{T}})](\lambda c.c\bm{\mathrm{10}}))\bm{\mathrm{F}}\\
&= [\lambda pc.c(\bm{\mathrm{S}}(p\bm{\mathrm{T}}))(p\bm{\mathrm{T}})][\lambda c.c\bm{\mathrm{21}}]\bm{\mathrm{F}}\\
&= [\lambda c.c\bm{\mathrm{32}} ])\bm{\mathrm{F}}\\
&= \bm{\mathrm{2}}
\end{aligned}
\end{equation*}

Of course now the subtraction function can be found (using \textbf{D} for difference):
\begin{equation*}
\bm{\mathrm{D}} = \lambda ab.b\bm{\mathrm{P}}a
\end{equation*}
Bearing in mind the predecessor of \textbf{0} is \textbf{0}, this will produce \textbf{0} if $b>a$.
\subsubsection{Further Conditional Logic}

@@

\begin{equation*}
\bm{\mathrm{R}} = \lambda y.(\lambda x.y(xx))(\lambda x.y(xx))
\end{equation*}
\begin{equation*}
\begin{aligned}
\bm{\mathrm{R}}a &= (\lambda x.a(xx))[\lambda x.a(xx)]\\
&= a([\lambda x.a(xx)][\lambda x.a(xx)])\\
&= a(\bm{\mathrm{R}}a)
\end{aligned}
\end{equation*}
Therefore, the \textbf{R} function applied to any function $a$ applies the function $a$ to the function \textbf{R} applied to $a$ - if substitution continues, an infinite series of nested $a$ functions is created; an infinite result is produced from a finite configuration. Of course we can create finites as well by halting the function - an example of this would be the factorial function, where \textbf{!}a = \textbf{M}a(\textbf{!}[\textbf{P}a]) and \textbf{!0} = \textbf{1}. Firstly, a recursive definition:
\begin{equation*}
\begin{aligned}
\bm{\mathrm{\xi}} &= \lambda fa.\bm{\mathrm{Z}}a\bm{\mathrm{1}}[\bm{\mathrm{M}}a(f[\bm{\mathrm{P}}a])]\\
\bm{!} &= \bm{\xi}!
\end{aligned}
\end{equation*}
i.e. the factorial function is the $\bm{\xi}$ function with itself as the argument $f$. Unfortunately such a definition is illegal under $\lambda$-Calculus as strictly no functions have names - this is where the function \textbf{R} comes into play. To find the factorial of \textbf{3}:
\begin{equation*}
\begin{aligned}
\bm{\mathrm{R\xi 3}} &= \bm{\mathrm{\xi}}(\bm{\mathrm{R\xi}})\bm{\mathrm{3}}\\
&= \bm{\mathrm{Z31}}[\bm{\mathrm{M3}}([\bm{\mathrm{R\xi}}][\bm{\mathrm{P3}}])]\\
&= \bm{\mathrm{M3}}(\bm{\mathrm{R\xi}}\bm{2})\\
&= \bm{\mathrm{M3}}(\bm{\mathrm{M2}}(\bm{\mathrm{M11}}))\\
&= \bm{\mathrm{6}}
\end{aligned}
\end{equation*}
The recursive identity \textbf{R} allows for \textbf{R}$\bm{\xi}$ to be applied as the first argument to the $\bm{\xi}$ function - as a result, the procedure also processes \textbf{2} and \textbf{1}, before stopping at \textbf{0} (which produces simply \textbf{1}. To simplify, the factorial function:
\begin{equation*}
\bm{!} = \lambda x.\bm{\mathrm{R\xi}}x
\end{equation*}
